### 2019-10-19 abstract/paper.tex
abstract in progress

\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\usepackage{url}
\usepackage{hyperref}

\linespread{1.03}

\acmConference[LAFI'2020]{Languages for Inference}{January 21,
2020}{New Orleans, LA, USA}

\setcopyright{none}

\bibliographystyle{ACM-Reference-Format}

\title{Probabilistic Programming around Gaussian Processes}
\author{David Tolpin}
\affiliation{
    \institution{PUB+}
    \country{Israel}
}
\email{david.tolpin@gmail.com}

\begin{abstract}
	We present GoGP, a library for probabilistic programming around
	Gaussian processes. Kernels, beliefs about hyperparameters
	and about observation inputs and outputs are
	programmatically expressed using the same differentiable
	probabilistic programming framework. One one hand, a basic
	usage with maximum-likelihood hyperparameter estimates and
	homoscedastic Gaussian noise requires as little coding as
	with any Gaussian process library. On the other hand,
	imposing prior beliefs on hyperparameters, handling input
	uncertainty, heteroscedastic and non-Gaussian noise, or
	change point detection can be organically added on top of a
	Gaussian process kernel. Just like distributions in
	probabilistic programs, kernels can be flexibly combined
	using general code flow rather than through predefined
	combination operators (usually addition and multiplication).
	Implemented in Go, a general-purpose programming language 
	popular for server-side programming, our library is well
	suited for integration into data processing/machine learning
	pipelines.  We demonstrate different uses of the library on
	a variety of case studies inspired by applications.
\end{abstract}

\begin{document}
\maketitle

\begin{sloppypar}
	\section{Gaussian Processes in Applications}

	Gaussian processes~\cite{} possess properties that make
	them the approach of choice in many applications:
	\begin{itemize}
		\item A Gaussian process works with as little or as much
			data as available.
		\item Non-uniformly sampled observations, missing
			observations, and observation noise are handled
			organically.
		\item Libraries for inference with Gaussian processes
			are readily available, both standalone and within
			ubiquitous machine learning libraries, such that
			Gaussian processes can be included into data
			processing pipelines and combined with other machine
			learning algorithms.
	\end{itemize}

\section{Challenges of Server-Side Probabilistic Programming}

Incorporating a probabilistic program, or rather a probabilistic
procedure, within a larger code body appears to be rather
straightforward: one implements the model in the probabilistic
programming language, fetches and preprocesses the data in the
host programming language, passes the data and the model to an
inference algorithm, and post-processes the results in the
host programming language again to make algorithmic
decisions based on inference outcomes. However, complex
server-side software systems make integration of probabilistic
inference challenging.

\paragraph{Simulation vs. inference} Probabilistic models
often follow a design pattern of simulation-inference: a
significant part of the model is a simulator, running an
algorithm with fixed parameters; the optimal parameters, or
their distribution, are to be inferred. The inferred parameters
are then used by the software system to execute the simulation
independently of inference for forecasting and decision making.

This pattern suggests re-use of the simulator: instead of
implementing the simulator twice, in the probabilistic model and
in the host environment, the same code can serve both purposes.
However to achieve this, the host language must coincide with
the implementation language of the probabilistic model, on one
hand, and allow a computationally efficient implementation of
the simulation, on the other hand. Some probabilistic systems
(Figaro~\cite{P09}, Anglican~\cite{TMY+16}, Turing~\cite{GXG18})
are built with tight integration with the host environment in
mind; more often than not though the probabilistic code is
not trivial to re-use.

\paragraph{Data interface} In a server-side application data
for inference comes from a variety of sources: network,
databases, distributed file systems, and in many different
formats. Efficient inference depends on fast data access and
updating. Libraries for data access and manipulation are
available in the host environment. While the host environment
can be used as a proxy retrieving and transforming the data,
such as in the case of Stan~\cite{Stan17} integrations,
sometimes direct access from the probabilistic code is the
preferred option, for example when the data is streamed or
retrieved conditionally.

\paragraph{Integration and deployment} Deployment of server-side
software systems is a delicate process involving automatic
builds and maintenance of dependencies. Adding  a component,
which possibly introduces additional software dependencies or
even a separate runtime, complicates deployment. Minimizing the
burden of probabilistic programming on the integration
and deployment process should be a major consideration in
design or selection of probabilistic programming tools.
Probabilistic programming systems that are implemented or 
provide an interface in a popular programming language, e.g.
Python (Edward~\cite{THS+17}, Pyro~\cite{Pyro18}) are easier
to integrate and deploy, however the smaller the footprint
of a probabilistic system, the easier is the adoption.

\section{Probabilistic Programming Facility for Go}

Based on the experience of developing and deploying solutions
using different probabilistic environments, we came up with
guidelines to implementation of a probabilistic programming
facility for server-side applications. We believe that these
guidelines, when followed, help easier integration of
probabilistic programming inference into large-scale server-side
software systems.

\begin{enumerate}
\item A probabilistic model should be programmed in the host
programming language. The facility may impose a discipline on
model implementation, such as through interface constraints, but
otherwise supporting unrestricted use of the host language for
implementation of the model.

\item Built-in and user-defined data structures and libraries
should be accessible in the probabilistic programming model.
Inference techniques relying on the code structure, such as
those based on automatic differentiation, should support the
use of common data structures of the host language.

\item The model code should be reusable between inference and
simulation. The code which is not required solely for inference
should be written once for both inference of parameters and use
of the parameters in the host environment.  It should be
possible to run simulation outside the probabilistic model without
runtime or memory overhead imposed by inference needs.
\end{enumerate}

In line with the guidelines, we have implemented a probabilistic
programming facility for the Go programming language,
\texttt{infergo} (\url{http://infergo.ml/}). We have chosen Go
because Go is a small but expressive programming language with
efficient implementation, which has recently become quite
popular for computation-intensive server-side programming. At
the time of writing, the implementation is still at an early
stage. However, this facility is already used in production
environment for inference of mission-critical algorithm
parameters.

A probabilistic model in \texttt{infergo} is an implementation
of the \texttt{Model} interface requiring a single method
\texttt{Observe} which accepts a vector (a Go \textit{slice}) of
floats, the parameters to infer, and returns a single float,
interpreted as unnormalized log-likelihood of the posterior
distribution. Implementation of model methods can be written
in virtually unrestricted Go and use any Go libraries.

For inference, \texttt{infergo} relies on automatic
differentiation. The source code of the model is
translated by a command-line tool provided by \texttt{infergo}
into an equivalent model with reverse-mode automatic
differentiation of the log-likelihood with respect 
to the parameters applied. The differentiation operates
on the built-in floating-point type and incurs only a small
computational overhead. However, even this overhead is avoided
when the model code is executed outside of inference algorithms:
both the original and the differentiated model are
simultaneously available to the rest of the program code, so
the methods can be called on the differentiated model for
inference, and on the original model for the most efficient
execution with the inferred parameters.

The Go programming language and development environment offer
capabilities which made implementation of \texttt{infergo}
affordable.

\begin{itemize}
    \item The Go parser and abstract syntax tree serializer are
        a part of the standard library. Parsing, transforming,
        and generating Go source code is straightforward and
        effortless.
	\item Type inference (or \textit{type checking} as it is
		called in the Go ecosystem), also provided in the
		standard library, augments parsing and allows to
		selectively apply transformation-based automatic
		differentiation  based on static expression types. 
	\item Go compiles and runs fast. Fast compilation and
		execution speeds allow to use the same facility for both
		exploratory design of probabilistic models and for
		inference in production environment.
	\item Go offers efficient parallel execution as a
		first-class feature, via so-called \textit{goroutines}.
		Goroutines streamline implementation of sampling-based
		inference algorithms. Sample generators and consumers
		are run in parallel, communicating through channels. 
		Inference is easy to parallelize in order to exploit
		hardware multi-processing, and samples are retrieved
		lazily for postprocessing. 
\end{itemize}

Table~\ref{tab:memory-runtime} provides memory consumption and
running time measurements on basic models to
illustrate  \texttt{infergo}'s performance.
The measurements were obtained on a 2.3GHz Intel Core 5
CPU with 8GB of memory for 1000 iterations of Hamiltonian Monte
Carlo with 10 leapfrog steps. Note that log-likelihood computation 
for standard distributions is not optimized yet. Quite the
opposite: since models in \texttt{infergo} are fully composable,
primitive distributions are themselves implemented as
\texttt{infergo} models and automatically differentiated.

{\smaller
\begin{table}[H]
	\caption{Memory and running times for 1000 iterations
	of HMC with 10 leapfrog steps.}
	\label{tab:memory-runtime}
	\begin{tabular}{r | c |  c | c}
		{\it model}  & \multicolumn{2}{c|}{\it time} & {\it memory} \\ 
		 & {\it compilation} & {\it execution} & \\\hline
		8 schools & 0.15s & 0.6s & 5.5MB \\
		10D normal, 100 points & 0.15s & 3.0s & 5.7MB \\
		50D normal, 100 points & 0.15s & 13.0s & 5.8MB 
	\end{tabular}
\end{table}}

A lightweight probabilistic programming facility similar to
\texttt{infergo} can be added to most modern general-purpose
programming languages, in particular those used in implementing
large-scale software systems, making probabilistic
programming inference more accessible in server-side
applications.
\end{sloppypar}

\bibliography{refs}

\end{document}